---
title: "exp4_pulpPhosphoDeconvolution"
output: html_document
author: Sara Gosline
date: "2023-02-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(ggfortify)
library(cowplot)
#library(leapR)
library(dplyr)
library(pheatmap)

source('spleenDataFormatting.R')
source('spatialProtUtils.R')

```

The assignment of pulp type from the basic data was not robust. Here we will use the cell type signatures from the sorted data to label the voxels.


## Voxel assignments from bulk data


```{r cell type signatures}
#run diffex for sorted prot
fullmap<-scater::runPCA(global.sorted)

full<-spatialDiffEx(fullmap,'pulpAnnotation')%>%
  rowData(.)%>%
  as.data.frame()%>%
  dplyr::select(logFC='pulpAnnotation.limma.logFC',adj.P.Val='pulpAnnotation.limma.adj.P.Val',
                AveExpr='pulpAnnotation.limma.AveExpr')

pumap<-scater::runPCA(spat.prot)



whitesig<-full%>%
  subset(adj.P.Val<0.01)%>%
  subset(logFC>1)%>%
  subset(AveExpr>1)

redsig<-full%>%
  subset(adj.P.Val<0.01)%>%
  subset(logFC<(-1))%>%
  subset(AveExpr>(1))

#score spatial prot
pumap<-calcSigScore(pumap,rownames(redsig),'RedPulp')%>%
  calcSigScore(rownames(whitesig),'WhitePulp')

#assign signature scores
newDat<-colData(pumap)%>%
  as.data.frame()%>%
  mutate(isRed=RedPulp>0.5,isWhite=WhitePulp>0.5)%>%
  mutate(pulp=ifelse(isRed,'red',ifelse(isWhite,'white','None')))


#map signatures to spatial data
spat.phos<-adjustPhophoWithGlobal(spat.phos,pumap)
spat.phos<-scater::runPCA(spat.phos)
colData(spat.phos)<-DataFrame(newDat)
p<-plotPCA(spat.phos,colour_by='pulp')
p
ggsave("phosphoPCAlabeled.png",p)
```

## Differential expression

Now we can use the functions we defined to do differential expression on the phosphosites, as well as ordered enrichment.

```{r diffex, echo=FALSE}

spat.phos<-spatialDiffEx(spat.phos,column='pulp',vals=c('red','white'))

res<- spat.phos%>%
  rowData(.)%>%
  as.data.frame()%>%
  dplyr::select(featureID='X',
                logFC='pulp.limma.logFC',
                adj.P.Val='pulp.limma.adj.P.Val',
                AveExpr='pulp.limma.AveExpr')


print(head(res))
res<-res%>%mutate(signif=(adj.P.Val<0.01)&(abs(logFC)>1))

num.diff<-nrow(subset(res,signif))
               
p<-res%>%
  mutate(logPval=log10(adj.P.Val)*-1)%>%
  ggplot(aes(x=logFC,y=logPval,col=signif))+geom_point()+
  ggtitle(paste0(num.diff,' differentially expressed phosphosites'))

p

ggsave('diffExPsites.png',p)

pheatmap(exprs(spat.phos)[rownames(res[which(res$signif),]),],
          annotation_col = as.data.frame(colData(spat.phos)[,c('RedPulp','WhitePulp')]),
         clustering_distance_cols = 'correlation',
         clustering_distance_rows = 'correlation',filename='diffExPsitesheatemap.png')
         
```

These signatures cluster voxels, but don't serve as great markers because they are still very much 'down' regulated. We want to pick the top expressed proteins that represent each pulp type.

```{r reduced signature}

library(leapR)
data('kinasesubstrates')

ksite<-res%>%
  tidyr::separate(featureID,into=c('gene','site'))%>%
  tidyr::unite(c(gene,site),col='psite',sep='-')%>%
  arrange(logFC)
                    
kres<-leapR::leapR(kinasesubstrates,enrichment_method='enrichment_in_order',
                   ksite,id_column='psite',primary_columns='logFC')

sig.kin<-subset(kres,BH_pvalue<0.05)%>%
  tibble::rownames_to_column('Kinase')

p1<-ggplot(sig.kin,aes(x=zscore,y=Kinase))+geom_bar(stat='identity')

p2<-ggplot(sig.kin,aes(x=log10(BH_pvalue)*-1,y=Kinase))+geom_bar(stat='identity')

p3<-cowplot::plot_grid(p1,p2)
p3
ggsave('Kinaseenrichment.png',p3)


```

So we have a handful of proteins that are marking the signature. Let's see if we can link them in a biological fashion.

```{r significant kniase substrate plotting}
#library(clusterProfiler)
library(pheatmap)

substr<-sig.kin%>%
  dplyr::select(Kinase,ingroupnames)%>%
  tidyr::separate(ingroupnames,into=c('s1','s2','s3','s4','s5','s6','s7','s8','s9','s10','s11','s12'),sep=', ')%>%
  pivot_longer(cols=starts_with('s'),names_to='sval',values_to='substrate')%>%
  subset(!is.na(substrate))%>%
  tidyr::separate(substrate,into=c('gene','mod'))%>%
  tidyr::unite(gene,mod,col='substrate',sep='_')

pheatmap(exprs(spat.phos)[unique(substr$substrate),],
         annotation_col = as.data.frame(colData(spat.phos)[,c('RedPulp','WhitePulp')]),
         clustering_distance_cols = 'correlation',
         clustering_distance_rows = 'correlation')


pheatmap(exprs(spat.phos)[unique(substr$substrate),],
         annotation_col = as.data.frame(colData(spat.phos)[,c('RedPulp','WhitePulp')]),
         clustering_distance_cols = 'correlation',
         clustering_distance_rows = 'correlation',
         filename = 'substratesOfEnrichmdKinaseHeatmap.pdf')
```

This isn't quite right - the 'green' items are definitely on the right, but we need some sort of threshold to 'call' the white vs. red pulp items.


### Plot in image

Lastly we can plot each set of kinase scores in an image by taking the signature/ranking signature approach we did for the bulk data. 

```{r plot image}

all.res<-lapply(unique(substr$Kinase),function(x){
  subs<-subset(substr,Kinase==x)$substrate
  res<-calcSigScore(spat.phos,subs,x)
  plotSigGrid(res,x)
})
  
k1<-all.res[[1]]
k2<-plotSigGrid(spat.phos,'WhitePulp')
 
kres<-cowplot::plot_grid(k1,k2)

kres

ggsave('csnk2a1res.png',kres)
```

It doesn't really seem like CSNK2A1 kinase activity is that correlated with white pulp from this image. 

What if we just plot the scores?

```{r kinase scoring by voxel}
vdat<-colData(spat.phos)%>%
  as.data.frame()%>%
  tibble::rownames_to_column('Voxel')

tab<-expToLongForm(spat.phos,'substrate')%>%
  right_join(substr)

tab%>%left_join(vdat)%>%ggplot(aes(x=pulp,y=LogRatio,col=pulp))+
  geom_jitter()+
  facet_grid(~Kinase)

p<-tab%>%left_join(vdat)%>%
  subset(pulp!='None')%>%
  ggplot(aes(x=pulp,y=LogRatio,col=pulp))+
  geom_boxplot()+
  facet_grid(~Kinase)+
  theme_bw()

p

ggsave('kinaseSubstrateDist.png',p)

```


All the expression of the substrates is pretty low - maybe we can try to find another way to bring in kinases associated with
white pulp? 

## Correlation based kinase enrichment

Here we will score substrates with their correlation to the white pulp score. 

```{r correlation scoring }

spat.phos<-calcCorrelationWithScore(spat.phos,'WhitePulp')

cor.white<-rowData(spat.phos)%>%
    as.data.frame()%>%
    tidyr::separate(X,into=c('gene','mod'),sep='_')%>%
    tidyr::unite(gene,mod,col='subs',sep='-')%>%
    leapR::leapR(kinasesubstrates,
               enrichment_method = 'enrichment_in_order',
               datamatrix=.,id_column='subs',primary_columns='WhitePulp.correlation')%>%
  subset(!is.na(pvalue))%>%
  arrange(pvalue)
             
print(head(cor.white))
spat.phos<-calcCorrelationWithScore(spat.phos,'RedPulp')

cor.red<-rowData(spat.phos)%>%
    as.data.frame()%>%
    tidyr::separate(X,into=c('gene','mod'),sep='_')%>%
    tidyr::unite(gene,mod,col='subs',sep='-')%>%
    leapR::leapR(kinasesubstrates,
               enrichment_method = 'enrichment_in_order',
               datamatrix=.,id_column='subs',primary_columns='RedPulp.correlation')%>%
  subset(!is.na(pvalue))%>%
  arrange(pvalue)

print(head(cor.red))

```

This is not super productive, nothing is statistically significant after FDR correction. So, we circle back to our network analysis.

```{r network}

diffex.net<-buildNetwork(spat.phos,featName='pulp.limma.logFC')
plot.PCSF(diffex.net$graph)


```