---
title: "Figure 3_4 histology enrichment"
output: html_document
author: Sara Gosline
date: "2023-07-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(ggfortify)
library(cowplot)
#library(leapR)
library(dplyr)

source('spleenDataFormatting.R')
#library(spammer)
source('spatialProtUtils.R')
```

The assignment of pulp type from the basic data was not robust. Here we will use the cell type signatures from the sorted data to label the voxels.


## Cell type signatures

### using sorted cells
Here we get the differential expression from the sorted cells, and try to map them to spatial

```{r cell type signatures}


##get fit values for diffex
pumap<-scater::runPCA(spat.prot)

phumap<-spat.phos%>%
  scater::runPCA()

pumap<-spatialDiffEx(pumap,column='Histology',vals=c('red','white'),'Protein')
phumap<-spatialDiffEx(phumap,column='Histology',vals=c('red','white'),'X')

```


Now we have loaded the data and can do enrichment

```{r prot enrich}
library(leapR)
library(org.Hs.eg.db)
data('krbpaths')
#map<-as.list(org.Hs.egSYMBOL2EG)ß
#whitegenes<-lapply(rownames(whitesig),function(x) return(map[[x]][1]))
#universe<-unlist(lapply(rownames(full),function(x) return(map[[x]][1])))

full<-rowData(pumap)|>
  as.data.frame()



prot.enrich<-leapR::leapR(geneset=krbpaths,
                         enrichment_method='enrichment_in_sets',id_column='Protein',
                    datamatrix=full,primary_columns='Histology.limma.logFC',greaterthan=T,threshold=0.25)

sig.enrich<-prot.enrich%>%
  subset(pvalue<0.01)|>
  arrange(pvalue)

plotResult(sig.enrich)
kr<-plotResult(sig.enrich)
kr
ggsave('fig3a_keggenrich.pdf',kr)
gosigs <- leapR::read_gene_sets('GO_Biological_Process_2021.txt')

go.enrich<-leapR::leapR(geneset=gosigs,
                         enrichment_method='enrichment_in_sets',id_column='Protein',
                    datamatrix=full,primary_columns='Histology.limma.logFC',greaterthan=T,threshold=0.2)|>
  subset(ingroup_n>1)

sig.go<-go.enrich%>%
  subset(pvalue<0.01)|>
  arrange(pvalue)

gr<-plotResult(sig.go)
gr

ggsave('fig3b_goEnrich.pdf',gr)

##plot GO Enrichment
prots<-c('CR2','MS4A1')

allfigs<-lapply(prots,function(x) plotFeatureGrid(pumap,x,x,'Histology'))
##plot heterochromatin

##plot snoRNA localizaton

##plot CR2 and MS4A1 expression
prot_plot=cowplot::plot_grid(plotlist=allfigs,labels=prots,ncol=1)
prot_plot
ggsave('fig3bd_protPlot.pdf',prot_plot)
```
```{r cell type}


protDiff<- full|>
  dplyr::select(featureID='Protein',
                logFC='Histology.limma.logFC',
                adj.P.Val='Histology.limma.adj.P.Val',
                AveExpr='Histology.limma.AveExpr')
####CAN we use MCP counter?
mcptab<-read.table('mcp-genelist.txt',sep='\t',header=T,check.names = F)
mcplist<-lapply(unique(mcptab$`Cell population`),function(x) 
  paste(c(x,mcptab$`HUGO symbols`[which(mcptab$`Cell population`==x)]),sep='\t'))
maxlength=max(sapply(mcplist,length))

mcpmat<-do.call(rbind,
                lapply(mcplist,function(x) paste(c(x,rep("",maxlength-length(x))),sep='\t')))
mcp.gl=list(names=unique(mcptab$`Cell population`),desc="",sizes=sapply(mcplist,length),matrix=mcpmat)
class(mcp.gl)<-c('geneset_data','list')

mcp.enrich<-leapR::leapR(geneset=mcp.gl,enrichment_method='enrichment_in_sets',id_column='featureID',
                        datamatrix=protDiff,primary_columns='logFC',greaterthan=T,threshold=0)|>
  subset(ingroup_n>1)

print(mcp.enrich)

###now try another matrix, LM22
##LM22 is not unique so not great
# lm22<-read.table('LM22.txt',sep='\t',header=T,check.names = F)|>
#    tidyr::pivot_longer(cols=c(2:23),names_to='cellType',values_to='weight')
# 
# lm22=lm22|> group_by(`Gene symbol`)|>
#    summarize(maxVal=max(weight),medVal=median(weight))|>
#   right_join(lm22)
# 
# dlm22<-lm22|>
#   mutate(overMed=weight<medVal,max=weight==maxVal)|>
#   subset(max)
# 
# lmlist<-lapply(unique(dlm22$cellType),
#                function(x) c(x,dlm22$`Gene symbol`[which(dlm22$cellType==x)]))
# lmsize<-sapply(lmlist,length)
# maxlength=max(lmsize)
# 
# lmmat<-do.call(rbind,
#                 lapply(lmlist,function(x) paste(c(x,rep("",maxlength-length(x))),sep='\t')))
# lm22.gl=list(names=unique(dlm22$cellType),desc="",sizes=sapply(lmlist,length),matrix=lmmat)
# class(lm22.gl)<-c('geneset_data','list')
# 
# lm22.enrich<-leapR::leapR(geneset=lm22.gl,enrichment_method='enrichment_in_sets',id_column='featureID',
#                         datamatrix=protDiff,primary_columns='logFC',greaterthan=T,threshold=0)|>
#   subset(ingroup_n>1)
# 
# print(lm22.enrich)

```

So we can find insignifincant enrichment in B celland T cells due to low coverage of markers. what if we just plot al markers?

```{r immune markers}
#now many are in the dataset!
#dlm22<-dlm22|>mutate(inProt=`Gene symbol`%in%rownames(exprs(pumap)))
rlm22<-lm22|>mutate(inProt=`Gene symbol`%in%rownames(exprs(pumap)))

print(subset(rlm22,inProt)|>group_by(cellType)|>summarize(n()))
##only 28 markers of the 547 are expressed in this dataset

expLM22<-subset(rlm22,inProt)
pheatmap(as.matrix(exprs(pumap)[intersect(expLM22$`Gene symbol`,rownames(rowData(pumap))),]),
         clustering_distance_cols = 'correlation',clustering_distance_rows='correlation',
         clustering_method = 'ward.D2',
         annotation_col = as.data.frame(colData(pumap))[,c('Histology','TMT.set')],
         main='White pulp upregulated',cellheight = 10,filename='fig3lm22Markers.pdf')

#figs<-lapply(unique(expLM22$`cellType`),function(ctype){
#  print(ctype)
#  prots<-expLM22$`Gene symbol`[which(expLM22$`cellType`==ctype)]
#  print(prots)
#  plotFeatureGrid(pumap,prots,gsub(" ",'',ctype),'Histology')
#})

#imp<-cowplot::plot_grid(plotlist=figs,labels=unique(expLM22$`cellType`),nrow=2)

#ggsave('fig3_lm22_mmuneMarkers.pdf',imp,width=12)
###mcpcounter better?
expMcp<-mcptab|>
  mutate(inProt=`HUGO symbols`%in%rownames(exprs(pumap)))|>
  subset(inProt)

###11 out of 111 mcpcounter markers are present
figs<-lapply(unique(expMcp$`Cell population`),function(ctype){
  print(ctype)
  prots<-expMcp$`HUGO symbols`[which(expMcp$`Cell population`==ctype)]
  plotFeatureGrid(pumap,prots,gsub(" ",'',ctype),'Histology')
})

imp<-cowplot::plot_grid(plotlist=figs,labels=unique(expMcp$`Cell population`),nrow=2)
ggsave('fig3_mcp_mmuneMarkers.pdf',imp,width=12)
```

The differential expression of the protein is not super significant

```{r phos enrich}
data('krbpaths')
data('kinasesubstrates')

phosDiff<-phumap|>rowData(.)%>%
  as.data.frame()%>%
  dplyr::select(featureID='X',
                logFC='Histology.limma.logFC',
                adj.P.Val='Histology.limma.adj.P.Val',
                AveExpr='Histology.limma.AveExpr')|>
  mutate(Phosphosite=stringr::str_replace(featureID,'_','-'))
  
pfDiff<-phosDiff|>
  tidyr::separate(featureID,into=c('prot','psite'))|>
  group_by(prot)|>
  summarize(mlfc=mean(logFC),medLfc=median(logFC),meanp=log(mean(adj.P.Val))*-1)

path.enrich<-leapR::leapR(geneset=krbpaths,
                          enrichment_method='enrichment_in_sets',
                          datamatrix=as.data.frame(pfDiff),
                          id_column='prot',primary_columns='mlfc',greaterthan=T,threshold=.45)|>
  subset(pvalue<0.05)

plotResult(path.enrich)

gosigs <- leapR::read_gene_sets('GO_Biological_Process_2021.txt')

go.enrich<-leapR::leapR(geneset=gosigs,
                         enrichment_method='enrichment_in_sets',id_column='prot',
                    datamatrix=as.data.frame(pfDiff),primary_columns='mlfc',greaterthan=T,threshold=0.45)|>
  subset(ingroup_n>1)|>
  subset(pvalue<0.05)

plotResult(go.enrich)

kin.enrich<-leapR::leapR(geneset=kinasesubstrates,
                         enrichment_method='enrichment_in_sets',
                         datamatrix=phosDiff,
                         id_column='Phosphosite',primary_columns='logFC',greaterthan=T,threshold=.45)|>
    subset(ingroup_n>1)|>
  subset(pvalue<0.05)

kin<-plotResult(kin.enrich)
kin
ggsave('fig4a_kinaseEnrichment.pdf',kin)

csnk2a1<-stringr::str_split(kin.enrich['CSNK2A1','ingroupnames'],', ')|>unlist()

p1<-lapply(csnk2a1,function(x) plotFeatureGrid(phumap,gsub('-','_',x),x,'Histology'))
p1<-cowplot::plot_grid(plotlist=p1,nrow=4,labels=csnk2a1)
ggsave('fig4b_sigSubs.pdf',p1,width=12)

```



### Plot in image

Lastly we want to plot a set of scores, in this case the signature scores, in an image. 


```{r plot image}


 p <- plotSigGrid(pumap,'RedPulp')
 p1<-plotSigGrid(pumap,'WhitePulp')
    

 pc<-cowplot::plot_grid(p,p1)
 pc
 ggsave('fig2_scoredImage.pdf',pc,width=14,height=4)
```

Now that we have a rank score, we can "call" each voxel. 



