---
title: "Figure 3 Spatial Signatures"
output: html_document
date: "2023-05-05"
---

The goal of this MD is to evaluate the combination of proteomic and phosphoproteomic measurements to evaluate the spleen data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("spleenDataFormatting.R")
source('spatialProtUtils.R')
library(pheatmap)
```

## First collect and annotate voxels by signature

First lets combine the signatures to see what we get


```{r combined}
pumap<-scater::runPCA(spat.prot)


##generate two phospho analyses
phumap<-spat.phos%>%
  scater::runPCA()

adj.phumap<-spat.phos%>%
  adjustPhophoWithGlobal(pumap)|>## do we subtract out global? or not? 
  scater::runPCA()


fullmap<-scater::runPCA(global.sorted)
fullmap<-spatialDiffEx(fullmap)

full<- fullmap%>%
  rowData(.)%>%
  as.data.frame()%>%
  dplyr::select(featureID='X',
                logFC='pulpAnnotation.limma.logFC',
                adj.P.Val='pulpAnnotation.limma.adj.P.Val',
                AveExpr='pulpAnnotation.limma.AveExpr')

upsig<-full%>%
  subset(adj.P.Val<0.01)%>%
  subset(logFC>1)|>  subset(AveExpr>(1))


downsig<-full%>%
  subset(adj.P.Val<0.01)%>%
  subset(logFC<(-1))|>
    subset(AveExpr>(1))

pumap<-calcSigScore(pumap,rownames(downsig),'RedPulp')%>%
  calcSigScore(rownames(upsig),'WhitePulp')

colData(phumap)[['RedPulp']]<-colData(pumap)$RedPulp
colData(phumap)[['WhitePulp']]<-colData(pumap)$WhitePulp


colData(adj.phumap)[['RedPulp']]<-colData(pumap)$RedPulp
colData(adj.phumap)[['WhitePulp']]<-colData(pumap)$WhitePulp

```
## Basic diffex and functional enrichment

Based on annotated voxels, let's compute differentially expressed proteins and phosphosites and store in table. 


```{r fdifferential expression}

newVals<-colData(pumap)%>%
  as.data.frame()%>%
  mutate(isRed=RedPulp>0.5,isWhite=WhitePulp>0.5)%>%
  mutate(pulp=ifelse(isRed,'red',ifelse(isWhite,'white','None')))

colData(pumap)[['pulp']]<-newVals$pulp
colData(phumap)[['pulp']]<-newVals$pulp
colData(adj.phumap)[['pulp']]<-newVals$pulp

protDiff<-spatialDiffEx(pumap,column='pulp',feat='Protein')|>
  rowData(.)%>%
  as.data.frame()%>%
  dplyr::select(featureID='Protein',
                logFC='pulp.limma.logFC',
                adj.P.Val='pulp.limma.adj.P.Val',
                AveExpr='pulp.limma.AveExpr')

sigProts<-subset(protDiff,adj.P.Val<0.05)|>subset(logFC>.5)|>
  arrange(logFC)

phosDiff<-spatialDiffEx(phumap,column='pulp',feat='X')|>
  rowData(.)%>%
  as.data.frame()%>%
  dplyr::select(featureID='X',
                logFC='pulp.limma.logFC',
                adj.P.Val='pulp.limma.adj.P.Val',
                AveExpr='pulp.limma.AveExpr')|>
  mutate(Phosphosite=stringr::str_replace(featureID,'_','-'))

sigPhos<-subset(phosDiff,adj.P.Val<0.01)|>subset(logFC>1)|>
  arrange(logFC)

pheatmap(as.matrix(exprs(phumap)[intersect(rownames(sigPhos),rownames(rowData(phumap))),]),
         annotation_col = as.data.frame(colData(phumap))[,c('row','pulp')],
         main='White pulp upregulated',cellheight = 10,filename='suppFullp0.05logfc1whitePhos.pdf')

aphosDiff<-spatialDiffEx(adj.phumap,column='pulp',feat='X')|>
  rowData(.)%>%
  as.data.frame()%>%
  dplyr::select(featureID='X',
                logFC='pulp.limma.logFC',
                adj.P.Val='pulp.limma.adj.P.Val',
                AveExpr='pulp.limma.AveExpr')|>
  mutate(Phosphosite=stringr::str_replace(featureID,'_','-'))

asigPhos<-subset(phosDiff,adj.P.Val<0.01)|>subset(logFC>1)|>
  arrange(logFC)

pheatmap(as.matrix(exprs(adj.phumap)[intersect(rownames(sigPhos),rownames(rowData(adj.phumap))),]),
         annotation_col = as.data.frame(colData(adj.phumap))[,c('row','pulp')],
         main='White pulp upregulated',cellheight = 10,filename='suppFullp0.05logfc1CorrectedwhitePhos.pdf')

###let's do heatmaps of these proteins 


```

There are still numerous differentially expressed phosphosites accounting for protein levels. 

## Now we can look for biological pathways

```{r enrichment}
library(leapR)
library(org.Hs.eg.db)
data('krbpaths')
data('kinasesubstrates')


kin.enrich<-leapR::leapR(geneset=kinasesubstrates,
                         enrichment_method='enrichment_in_sets',
                         datamatrix=phosDiff,
                         id_column='Phosphosite',primary_columns='logFC',greaterthan=T,threshold=1)|>
    subset(ingroup_n>1)

sig.kin<-kin.enrich|>subset(pvalue<0.05)

kin.enrich<-leapR::leapR(geneset=kinasesubstrates,
                         enrichment_method='enrichment_in_sets',
                         datamatrix=aphosDiff,
                         id_column='Phosphosite',primary_columns='logFC',greaterthan=T,threshold=1)|>
    subset(ingroup_n>1)

a.sig.kin<-kin.enrich|>subset(pvalue<0.05)



##figure out what to do with this?
```

however, there are no significantly enriched kinases.

## Plot enrichment

Now we need to write a function to plot enrichment status

```{r plot enrichment}

plotResult<-function(enrich_res){
###the columns dpeendon the output a bit
  library(ggplot2)
  
  odds<-enrich_res|>
    arrange(desc(oddsratio))
   
  ##get the top 20
  odds<-odds[1:min(20,nrow(enrich_res)),]|>
    tibble::rownames_to_column('Pathway')|>
    mutate(logPval=(-1*log10(pvalue)))
  
  odds$Pathway<-factor(odds$Pathway,levels=rev(odds$Pathway))
  
  ##plot
  res<-ggplot(odds,aes(x=Pathway,y=oddsratio,fill=logPval))+
    geom_bar(stat='identity')+
    coord_flip()

  res
}

kinplot<-plotResult(sig.kin)

```



## Kinase enrichment

We do not see a lot of coherent kinase activity, but there is some, we can investigate here.

```{r kinase enrichment and substrate ploting}

kinplot

ggsave('uncorrectedkinaseEnrichment.pdf',kinplot)


csnk2a1<-stringr::str_split(sig.kin['CSNK2A1','ingroupnames'],', ')|>unlist()

csnk2a2<-stringr::str_split(sig.kin['CSNK2A2','ingroupnames'],', ')|>unlist()

p1<-lapply(csnk2a1,function(x) plotFeatureGrid(phumap,gsub('-','_',x),x))

gp1<-cowplot::plot_grid(plotlist=p1,ncol=3)
gp1
ggsave('csnk2a1Subs.pdf',gp1,height=12,width=12)
p2<-lapply(csnk2a2,function(x) plotFeatureGrid(phumap,gsub('-','_',x),x))

gp2<-cowplot::plot_grid(plotlist=p2,ncol=2)
gp2

ggsave('csnk2a2Subs.pdf',gp2,height=6,width=8)


##what do they look like in normalized data?
p1<-lapply(csnk2a1,function(x) plotFeatureGrid(adj.phumap,gsub('-','_',x),x))
nulls<-which(unlist(lapply(p1,function(x) !is.null(x))))
p1<-p1[nulls]
gp1<-cowplot::plot_grid(plotlist=p1,ncol=3)
gp1
ggsave('adj.csnk2a1Subs.pdf',gp1,height=12,width=12)
p2<-lapply(csnk2a2,function(x) plotFeatureGrid(adj.phumap,gsub('-','_',x),x))
nulls<-which(unlist(lapply(p2,function(x) !is.null(x))))
p2<-p2[nulls]
gp2<-cowplot::plot_grid(plotlist=p2,ncol=2)
gp2

ggsave('adj.csnk2a2Subs.pdf',gp2,height=6,width=8)

```

## Networks show what we are missing with enrichment - mechanism? 


Let's try to build biological networks using the signatures.

```{r build networks,echo=F, warning=F}
library(PCSF)

whiteweights<-abs(sigProts$logFC)
names(whiteweights)<-sigProts$featureID

##add in phosphosites to PPI
whitephos<-abs(sigPhos$logFC)
names(whitephos)<-sigPhos$Phosphosite

fullnet<-buildNetwork(whiteweights,whitephos,beta=1,nrand=100,'sigPhosphoSites')

plot.PCSF(fullnet$graph)


whiteweights<-abs(asigPhos$logFC)
names(whiteweights)<-asigPhos$featureID

##add in phosphosites to PPI
whitephos<-abs(asigPhos$logFC)
names(whitephos)<-asigPhos$Phosphosite

fullnet<-buildNetwork(whiteweights,whitephos,beta=1,nrand=100,'sigCorrectedPhosphoSites')

plot.PCSF(fullnet$graph)


```
CD22 and CD19 are note measured, but implicated in the network. can we stain for them?


```